# Results

Our results begin with an exploration of key trends in Marvel media reception, providing context for the predictive modeling that follows. By first examining how critic and audience scores have evolved over time, we can see where the MCU has maintained its momentum and where reception has become more volatile. These patterns not only validate the motivation for our project but also highlight the value of looking to the comics themselves for insights. This section moves from broad descriptive trends to targeted model outputs, ultimately showing how specific narrative and structural traits align with adaptation success.

## Exploratory Data Analysis


To test our intuition that Marvel’s newer content has not been hitting the same highs as its earlier phases, we turned to the data. Figure 1 visualizes Rotten Tomatoes audience scores for Marvel media over time, from 2008 through 2025.


The trend is hard to miss: while early titles (roughly 2008–2018) generally received consistently high ratings—frequently in the 80–90% range—recent releases show a dramatic increase in variability. Starting around 2019, the scores became much more volatile, with several sharp dips into the 40s and 50s and far fewer guaranteed crowd-pleasers. Even well-established franchises are not immune to this drop in reception.


```{r echo = FALSE, message = FALSE}
#| fig-cap: "Figure 1: A time series graph showing the Rotten Tomatoes Audience Scores for Marvel Media over Time."
options(repos = c(CRAN = "https://cran.rstudio.com/"))
library(tidyverse)
marvel_ratings<-read_csv("C:/Users/lazyp/Downloads/marvel_ratings.csv")
ggplot(marvel_ratings, aes(x=release_date, y = audience_score))+
  geom_line()+
  geom_point()+
  geom_smooth(se=FALSE)+
  theme_minimal()+
  labs(title = "We're Not Crazy: Marvel's Just Not the Same",
       subtitle = "Marvel Media's Audience Score Over Time",
       x = "Release Date of Media",
       y = "Rotten Tomatoes Audience Score (Percent)",
       caption = "Data Source: Rotten Tomatoes")
```

These results back up what many longtime fans (ourselves included) have been feeling: the MCU just is not as consistent as it used to be. Audiences seem increasingly divided over the newer content, and not every experiment has paid off. At the same time, there has been a surge in the volume of media released. With more shows and movies hitting screens each year, the range of quality—and audience response—has grown.


We observed a similar pattern when looking at critic scores, as seen in Figure 3. While critic ratings historically showed less fluctuation than audience ratings, the post-2019 era still reveals a wider spread. Interestingly, the fitted trend line for critic scores slopes slightly upward in recent years, while the audience trend line slopes downward. This divergence suggests that critics and audiences may be responding differently to newer MCU content—a potential signal that what resonates with critics is not always what resonates with general audiences.
```{r echo=FALSE, message= FALSE}
#| fig-cap: "Figure 3: A time series graph showing the Rotten Tomatoes Critic Scores for Marvel Media over Time."
options(repos = c(CRAN = "https://cran.rstudio.com/"))
library(tidyverse)
marvel_ratings<-read_csv("C:/Users/lazyp/Downloads/marvel_ratings.csv")
ggplot(marvel_ratings, aes(x=release_date, y = critic_score))+
  geom_line()+
  geom_point()+
  geom_smooth(se=FALSE)+
  theme_minimal()+
  labs(title = "Not in Perfect Harmony: Critics vs. Crowds",
       subtitle = "Marvel Media's Critic Score Over Time",
       x = "Release Date of Media",
       y = "Rotten Tomatoes Critic Score (Percent)",
       caption = "Data Source: Rotten Tomatoes")
```

Together, these patterns confirm that our hunch was not just nostalgia: something has changed in how Marvel stories are being received. This increasing variability and the critic audience split reinforce our motivation to look deeper into the storytelling DNA of the comics themselves. If certain narrative traits in the source material consistently lead to strong reception, identifying them could help explain why some adaptations resonate while others falter.

## Model Performance and Insights
Following our exploratory analysis, we evaluated how well different modeling approaches could predict whether a comic would be a strong candidate for a successful MCU adaptation.


### Constrained XGBoost


The constrained XGBoost model emerged as the strongest binary classifier, achieving 89.9% accuracy, AUC of 0.968, Kappa of 0.4473, 89.4% sensitivity, and 92.9% specificity. This configuration effectively balanced high predictive performance with reduced overfitting.


Figure X shows the top 15 features ranked by “Gain,” XGBoost’s measure of how much each feature contributed to improving the model’s objective function. The most influential predictors were series_issue_count, theme_civil_conflict, num_creators, and page_count, suggesting that both the scope of a comic series and certain narrative conflicts are strong signals for adaptation success. Economic or market-related attributes such as price also ranked highly, hinting at potential links between production investment and adaptation likelihood. Additionally, thematic elements like weapons and supertech, romantic drama, and military conflict appeared in the top 15, reinforcing the importance of recognizable, high-stakes story elements in commercially viable adaptations.

```{r echo=FALSE, message= FALSE}
library(randomForest)
library(caret)
library(janitor)
df <- read_csv("C:/Users/lazyp/Downloads/fr_comics.csv") %>%
  clean_names()
  library(stringr)

df_imputed <- df %>%
  mutate(
    # Text fields
    comic_title = replace_na(comic_title, "Unknown"),
    comic_rating = replace_na(comic_rating, "Unknown"),

    # Numeric fields
    page_count = replace_na(page_count, 0),
    price = replace_na(price, 0),
    series_issue_count = replace_na(series_issue_count, 0),
    num_characters = replace_na(num_characters, 0),
    num_teams = replace_na(num_teams, 0),
    num_ethnicities = replace_na(num_ethnicities, 0),
    num_power_types = replace_na(num_power_types, 0),
    num_creators = replace_na(num_creators, 0),
    num_themes = replace_na(num_themes, 0),
    num_threats = replace_na(num_threats, 0),

    # Boolean flags → assume FALSE = not present if NA
    comedy = replace_na(comedy, FALSE),
    origin_story = replace_na(origin_story, FALSE),
    mentor = replace_na(mentor, FALSE),
    impactful_moment = replace_na(impactful_moment, FALSE),
    cultural = replace_na(cultural, FALSE),
    romance = replace_na(romance, FALSE),
    movie_based = replace_na(movie_based, FALSE),
    used_in_movie = replace_na(used_in_movie, FALSE),

    # Optional: release_date → leave as NA or impute
    release_date = replace_na(release_date, as.Date("1900-01-01"))
  )

library(fastDummies)
library(janitor)
library(tidyverse)
library(xgboost)
df_imputed <- df_imputed %>%
  mutate(across(where(~ is.logical(.) || all(. %in% c(TRUE, FALSE, NA))), ~ replace_na(., FALSE)))
  train_df <- df_imputed %>% filter(!is.na(success_score))
predict_df <- df_imputed %>% filter(is.na(success_score))

# Define binary label for success (can tune threshold here)
train_df <- train_df %>%
  mutate(success_label = ifelse(success_score >= 85, 1, 0)) %>%
  mutate(success_label = as.factor(success_label))
features <- train_df %>%
  select(-comic_id_1,-comic_id_25, -comic_id_120, -comic_title, -success_score, -release_date, -theme_cosmic_rivalries_power_struggles)
  features_xgb <- features%>% select(-series_id)
levels(features_xgb$success_label) <- c("neg", "pos")
ctrl <- trainControl(
  method = "cv",
  number = 5,
  classProbs = TRUE,
  summaryFunction = twoClassSummary,
  sampling = "up"  # you can also try "down" or "smote"
)

# Train constrained XGBoost
xgb_constrained <- train(
  success_label ~ .,
  data = features_xgb,
  method = "xgbTree",
  trControl = ctrl,
  metric = "ROC",
  tuneGrid = expand.grid(
    nrounds = 50,
    max_depth = 2,
    eta = 0.3,
    gamma = 5,
    colsample_bytree = 0.5,
    min_child_weight = 10,
    subsample = 0.7
  )
)

# Evaluate
xgb_preds <- predict(xgb_constrained, newdata = features_xgb)


# AUC
xgb_probs <- predict(xgb_constrained, newdata = features_xgb, type = "prob")[, "pos"]
xgb_imp <- xgb.importance(model = xgb_constrained$finalModel)

xgb_imp_df <- as.data.frame(xgb_imp)
ggplot(xgb_imp_df[1:15,], aes(x = reorder(Feature, Gain), y = Gain)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  theme_minimal()+
  labs(title = "XGBoosted: Top 15 Most Important Features", x = "Feature", y = "Gain", caption = "Data From Marvel.com")
```

### Logistic Regression


While slightly less accurate (81.2% accuracy, AUC of 0.898, 96.5% precision, 71.0% recall, Kappa of 0.425), logistic regression provided far greater interpretability. Coefficients revealed which traits were positively or negatively associated with adaptation success, allowing us to directly connect model outputs to storytelling elements.


As shown in Figure Y, the top coefficients by absolute value included theme_scientific_ethics, theme_civil_conflict, and threat_social_threat_uprising, indicating that moral dilemmas, internal societal conflicts, and large-scale uprisings are strongly tied to positive reception. Other notable features included theme_faith_and_religion, theme_coming_of_age, and theme_hero_vs_clone_doppelganger, suggesting that personal growth arcs and identity-based conflicts resonate with audiences when adapted.

```{r echo=FALSE, message= FALSE}
library(caret)
logistic_df <- features
levels(logistic_df$success_label) <- c("neg", "pos")
set.seed(42)
# Train logistic regression model
log_model <- train(
  success_label ~ .,
  data = logistic_df,
  method = "glm",
  family = "binomial",
  trControl = ctrl,
  metric = "ROC"
)

log_probs <- predict(log_model, newdata = logistic_df, type = "prob")[, "pos"]

library(broom)

# Extract coefficients and sort by absolute value
log_imp <- tidy(log_model$finalModel) %>%
  filter(term != "(Intercept)") %>%  # remove intercept
  mutate(abs_estimate = abs(estimate)) %>%
  arrange(desc(abs_estimate))

# Optional: visualize
library(ggplot2)
ggplot(log_imp[1:15,], aes(x = reorder(term, abs_estimate), y = abs_estimate)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 15 Logistic Regression Variables by Absolute Coefficient",
    x = "Feature",
    y = "Absolute Coefficient",
    caption = "Data from Marvel.com"
  ) +
  theme_minimal()
```
### Consensus Feature Importance
To combine predictive strength with interpretability, we averaged the probability outputs from both models to create a composite “success potential score.” Figure Z shows the consensus feature importance across both models. Traits like theme_cosmic_rivalries_power_struggles, theme_civil_conflict, and threat_national_emergency consistently ranked near the top, pointing to a shared recognition across models that large-scale stakes and high-intensity conflict drive adaptation success. Elements such as theme_infiltration_and_deception, theme_mystery_and_investigation, and theme_weapons_and_supertech also stood out, reflecting the enduring appeal of espionage, detective-style narratives, and advanced technology in the MCU formula.

```{r echo=FALSE, message= FALSE}
xgb_imp <- xgb.importance(model = xgb_constrained$finalModel) %>%
  rename(term = Feature) %>%
  mutate(importance_rank_xgb = rank(-Gain, ties.method = "first"))

# Already have log_imp from broom::tidy()
log_imp <- tidy(log_model$finalModel) %>%
  filter(term != "(Intercept)") %>%
  mutate(abs_estimate = abs(estimate),
         importance_rank_log = rank(-abs_estimate, ties.method = "first"))

# Now join successfully
combined_importance <- full_join(log_imp, xgb_imp, by = "term")

# Continue with your consensus ranking
consensus_importance <- combined_importance %>%
  mutate(
    importance_rank_log = ifelse(is.na(importance_rank_log),
                                 max(importance_rank_log, na.rm = TRUE) + 1,
                                 importance_rank_log),
    importance_rank_xgb = ifelse(is.na(importance_rank_xgb),
                                 max(importance_rank_xgb, na.rm = TRUE) + 1,
                                 importance_rank_xgb)
  ) %>%
  mutate(
    log_rank_norm = 1 - ((importance_rank_log - 1) / (max(importance_rank_log) - 1)),
    xgb_rank_norm = 1 - ((importance_rank_xgb - 1) / (max(importance_rank_xgb) - 1))
  ) %>%
  mutate(consensus_score = (log_rank_norm + xgb_rank_norm) / 2) %>%
  arrange(desc(consensus_score))

ggplot(head(consensus_importance, 15),
       aes(x = reorder(term, consensus_score), y = consensus_score)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Consensus Feature Importance\n(Logistic Regression + XGBoost)",
    x = "Feature",
    y = "Consensus Score (0–1)"
  ) +
  theme_minimal()

```

By combining model outputs and aligning them with top features from each, we produced not just a ranked list of comics with the highest adaptation potential, but also a clear thematic and structural profile of what “success” tends to look like in the MCU context.


## Final Model Recommendation
Our composite scoring system—an equal-weighted average of constrained XGBoost and logistic regression probabilities—ranked every comic in our dataset according to its potential for MCU adaptation success. While the model outputs a full ranked list, our top candidate was *Agents of Atlas (2006) #5*, with an XGBoost probability of 0.9988, a logistic regression probability of 1.0000, and a combined score of 0.9994.

This recommendation is supported by the presence of many traits that ranked highly in our consensus feature importance analysis (Figure Z). *Agents of Atlas (2006) #5* features rivalries and power struggles, civil conflict, and national emergency–level threats, all of which were among the top consensus predictors of adaptation success. It also incorporates strong elements of infiltration and deception and mystery and investigation, weaving espionage and covert operations into the plot. These thematic drivers, combined with the inclusion of advanced technology (“weapons and supertech”), a diverse team structure, and large-scale stakes, closely mirror the narrative DNA of past critically and commercially successful MCU adaptations.

It’s important to stress that this result does not suggest *Agents of Atlas (2006) #5* should be adapted in isolation. MCU projects almost always blend multiple comics and arcs into a single narrative. Instead, its high score reflects the fact that its narrative and structural features align strongly with those most associated with past MCU successes—making it an exemplary case study of what our models define as “adaptation-ready” storytelling.

